---
title: "p8105_hw3_ps3194"
author: "Pangsibo Shen"
date: "10/9/2020"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
data("instacart")
```

###### how many aisles, and which are most items from?

```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

###### Let's make a plot

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
###### Let's make a table!
```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients","dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

###### Apples vs ice cream.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

## Problem 2

###### load and tidy the data
```{r accel}
accel = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>%
  transform(
    minute = as.numeric(minute),
    #day = as_factor(day),
    week = as.factor(week)
  ) %>%
  mutate(weekday = case_when(
    day %in% c("Monday","Tuesday","Wednesday","Thursday","Friday") ~ "weekday",
    day %in% c("Saturday","Sunday") ~ "weekend")
  ) %>%
  transform(
    weekday = as.factor(weekday)
  )
 

head(accel)

```

The resulting dataset has 6 variables: week, day_id, day, minute, activity_count and weekday. The weekday variable has two levels: weekday and weekend. There are `r nrow(accel)` observations from the dataset.

###### relevel the day variable to chronological order
```{r relevel}
accel$day = factor(accel$day, c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
```

###### create a table showing the sum count for each day
```{r table}
accel %>%
  group_by(week, day_id, day) %>%
  summarise(
    sum_count = sum(activity_count)
  ) %>%
  arrange(week,day) %>%
  knitr::kable(align = 'c')
```
For the first weeks, the patient's daily total activity counts generally had increased from Monday to Sunday.For week 4, the patient's daily total activity counts had decreased from Monday to Saturday. For week 5, the patient's daily total activity counts had increased from Monday to Friday.

##### wrangling the date for the plot
```{r}
accel = accel %>%
  mutate(hour = case_when(
    minute %in% 1:60 ~ 1,
    minute %in% 61:120 ~ 2,
    minute %in% 121:180 ~ 3,
    minute %in% 181:240 ~ 4,
    minute %in% 241:300 ~ 5,
    minute %in% 301:360 ~ 6,
    minute %in% 361:420 ~ 7,
    minute %in% 421:480 ~ 8,
    minute %in% 481:540 ~ 9,
    minute %in% 541:600 ~ 10,
    minute %in% 601:660 ~ 11,
    minute %in% 661:720 ~ 12,
    minute %in% 721:800 ~ 13,
    minute %in% 801:860 ~ 14,
    minute %in% 861:920 ~ 15,
    minute %in% 921:980 ~ 16,
    minute %in% 981:1040 ~ 17,
    minute %in% 1041:1100 ~ 18,
    minute %in% 1101:1160 ~ 19,
    minute %in% 1161:1200 ~ 20,
    minute %in% 1201:1260 ~ 21,
    minute %in% 1261:1320 ~ 22,
    minute %in% 1321:1380 ~ 23,
    minute %in% 1381:1440 ~ 24,
    )) %>%
  group_by(week, day_id, day, hour) %>%
  summarise(
    hour_count = sum(activity_count)
  )

```
###### plot showing 24 hour activity
```{r}
accel %>%
  group_by(day, hour) %>%
  summarise(avg_hour_count = mean(hour_count)) %>%
  ggplot(aes(x = hour, y = avg_hour_count, color = day)) +
  geom_line(alpha = 10) +
  scale_x_continuous(breaks = accel$hour) + 
  ylab("avg hourly activity count by weeks") +
  guides(fill = guide_legend(title = "Week"))
```
This is a graph showing the 5 weeks average 24-hour activity count for each day in the week. From the graph, we noticed that the patient started to become active around 7 am (the patient wakes up). The activity count peaks around 13 pm and 21 pm and the peaks might be explained by exercising activities such as walking after lunch and before bed.   


## Problem 3

###### load dataset
```{r}
data("ny_noaa")
```

###### data cleaning and separate the date
```{r}
ny_noaa_tidy = ny_noaa %>%
  na.omit() %>%
  separate(date,
           c("year","month","day"),
           sep = "-") %>%
  transform(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    snow = as.character(snow)
  ) %>%
  mutate(
   prcp = prcp/10,
   tmax = tmax/10,
   tmin = tmin/10
  )

head(ny_noaa_tidy)
```
######  the most commonly observed values for snowfall
```{r}
ny_noaa_tidy %>%
  count(snow) %>%
  arrange(desc(n)) %>%
  head()
```
Other than days without snow (snowfall is 0), the most commonly observed value for snowfall is 25 mm.
 
```{r}
ny_noaa_tidy %>%
  filter(month %in% c("01","07")) %>%
  transform(year = as.numeric(year),
            id = as.factor(id)) %>%
  mutate(month = recode(month, "01" = "January", "07" = "July")) %>%
  group_by(id, year, month) %>%
  summarise(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = year, y = mean_tmax, group =id)) +
  geom_line() +
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, hjust = 1)) +
  facet_grid(cols = vars(month))
```

 
